---
title: "Statistical Inference Course Project Part 2"
author: "Sven Koerbitz"
output: pdf_document
---
##Overview
Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package. 

##1. Load the ToothGrowth data and perform some basic exploratory data analyses 
```{r}
library(datasets)
library(ggplot2)
data(ToothGrowth)
plot <- ggplot(ToothGrowth, aes(x=factor(dose),y=len,fill=factor(dose)))
plot + geom_boxplot() + 
     facet_grid(.~supp) +
     scale_x_discrete("Dose") +   
     scale_y_continuous("Length")  
```

##2. Provide a basic summary of the data.
```{r}
head(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
table(ToothGrowth$supp, ToothGrowth$dose)
```

##3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
```{r}
# T-Test by Supp Type
t.test(len ~ supp, data = ToothGrowth)

# Subsetting Data by Dose
dose0.5 = subset(ToothGrowth, dose == 0.5)
dose1.0 = subset(ToothGrowth, dose == 1.0)
dose2.0 = subset(ToothGrowth, dose == 2.0)

# T-Test Supp by Dose
t.test(len~supp, data=dose0.5)
t.test(len~supp, data=dose1.0)
t.test(len~supp, data=dose2.0)
```
Given a dose of 0.5 the null hypothesis that the supplement type (OJ/VC) has no influence on the growth of teeth can be rejected because the p-value is 0.006359 which is way below a significance level of 0.05.

Given a dose of 1.0 the null hypothesis that the supplement type (OJ/VC) has no influence on the growth of teeth can be rejected because the p-value is 0.001038 which is way below a significance level of 0.05.

Given a dose of 2.0 the null hypothesis that the supplement type (OJ/VC) has no influence on the growth of teeth can not be rejected because the p-value is 0.9639 which is above a significance level of 0.05. The data is not conclusive.

##4. State your conclusions and the assumptions needed for your conclusions. 
Based on the basic exploratory data analyses the assumption was made that
  - the growth of teeth depends on the dose (Higher dose = longer theeth) and
  - the supplement type OJ has a higher effect on the growth of teeth than the supplement type VC
  
Based on the hypothesis tests it can be concluded that 
  - the growth of teeth increases with the dose and
  - the supplement type OJ has a higher effect in the doses of 0.5 and 1.0 but not in doses of 2.0
